---
title: "Darwin Core mapping of Contrat de Rivière Senne data"
author:
- Peter Desmet
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

This document describes how (and contains the code to) transform Contrat de Rivière Senne data to a Darwin Core Archive that can be uploaded to an IPT.

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Install required libraries (if not yet installed):

```{r}
installed <- rownames(installed.packages())
required <- c("magrittr", "here", "glue", "readr", "DBI")
if (!all(required %in% installed)) {
  install.packages(required[!required %in% installed])
}
```

Load libraries:

```{r message = FALSE}
library(magrittr)       # To use pipes
library(here)           # To find files
library(glue)           # To insert variables in strings
library(readr)          # To write files
library(DBI)            # To create and query databases
```

# Read source data

Read source files:

```{r}
# All columns are read as characters for easier import into SQLite
data <- readr::read_csv(here::here("data", "raw", "inventaire_riparias.csv"), col_types = cols(.default = "c"))
```

## Create database

Create a SQLite database with the source data, so it can be queried with SQL in the next steps:

```{r}
db <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")

# Import data
DBI::dbWriteTable(db, "data", data)
```

## Darwin Core mapping

Create [Event](https://rs.gbif.org/core/dwc_event_2022-02-02.xml) core:

```{r}
dwc_event_sql <- glue::glue_sql(readr::read_file(here::here("sql", "dwc_event.sql")), .con = db)
dwc_event <- DBI::dbGetQuery(db, dwc_event_sql)
```

Create [Occurrence](https://rs.gbif.org/core/dwc_occurrence_2022-02-02.xml) extension:

```{r}
dwc_occurrence_sql <- glue::glue_sql(readr::read_file(here::here("sql", "dwc_occurrence.sql")), .con = db)
dwc_occurrence <- DBI::dbGetQuery(db, dwc_occurrence_sql)
```

## Save data to CSV

```{r}
write_csv(dwc_event, here::here("data", "processed", "event.csv"), na = "")
write_csv(dwc_occurrence, here::here("data", "processed", "occurrence.csv"), na = "")
```
